Evaluation of classical correlation functions from 2/3D images on CPU and GPU architectures: introducing CorrelationFunctions.jl

### 2022-10-04 paper-correlations.tex
Suddenly I must do another job

\section{Overview of supported functions and modules.}
\label{sec:overview}
\verb+CorrelationFunctions.jl+ is capable to compute the following correlation
functions:

These correlation functions are calculated using two different approaches (when
applicable). The first approach is implemented in the module \code{Directional}
and the second approach is in the module \code{Map}. The former approach (called
the scan approach) calculates correlation functions along specified directions
which can include axial and diagonal directions. The results of computations can
be provided to the user separately for each requested direction or averaged
across multiple directions. The latter approach (called the map approach)
calculates a ``correlation map'' which holds the values of a correlation
function in all directions. The scan and map approaches are described in
\cref{sec:scan} and \cref{sec:map} respectively.

Our package takes one-, two- or three-dimensional arrays of any element type
(usually we restrict ourselves to integer numbers of type \code{Int} or
\code{Int8}) and treats them as homogeneous anisotropic media. The package can
understand two-phase as well as multiphase systems. Indeed, a multiphase system
$A$ can be easily converted to a two-phase system $B$ by applying an indicator
function $\chi(x)$ to each element of $A$ which can be written as $B = \chi(A)$.

Correlation functions can be computed using periodic and zero-padding boundary
conditions. With periodic boundary conditions a correlation function is
calculated as if the input array were periodically extended to infinity,
i.e. $A_k = A_{k+N_k}$ where $N_k$ is the k-th array dimension. With
zero-padding the input is extended with zeros, i.e. $A_k = 0$ when
$0 < k < N_k$.

\section{The map approach.}
\label{sec:map}
In this approach we actively use fast Fourier transform (FFT)
to compute some correlation function ($S_2, F_{ss}, F_{sv}, C_2$)
in all directions.
The FFT algorithm let us compute auto and cross-correlation fast,
especially on GPU.

@@

\section{The scan approach.}
\label{sec:scan}
In this approach, the input array is cut in parallel slices (``scanned'') along
each requested direction. Each slice is then analyzed independently as
one-dimensional system and the result is accumulated into \code{CorrelationData}
structure. In the following sections we provide the algorithms for computation
of the functions given in \cref{sec:overview}.

>>>>

\verb+CorrelationFunctions.jl+ is capable to compute all these
functions. Usually there are two implementations for a function. The first
implementation computes a correlation function in specific predefined directions
(i.e. orientation of test vectors thrown into a sample is fixed). There are one
predefined direction in 1D case, four directions in 2D case and thirteen
directions in 3D case. This implementation is somewhat slow but requires much
less memory than the second one. The second implementation computes a
correlation function in all directions and is faster but requires much memory.
These implementations are contained in \code{Directional} and \code{Map} modules
of the package correspondingly.

These functions can work in periodic and non-periodic modes. Periodic mode
assumes that an input sample is periodically continued to the whole Euclidean
space. Non-periodic mode assumes zero-padded input.

\section{The map approach.}
\label{sec:map}
Definition \cref{eq:s2-def} can be rewritten as follows:
\begin{equation}
  S_2^{(i)}(r) = \frac{F^{-1} [|F [A^{(i)}](z)|] (r)}{N(r)} \label{eq:s2-ft}
\end{equation}
where $A^{(i)} \xleftarrow{I^{(i)}} A$ is a binary array obtained by element-wise
application of interface indicator $I^{(i)}$ to an input $A$, $F$ is a discrete
Fourier transform operator and $N(r)$ is a total number of trials (vectors
thrown into sample).

In this approach we actively use fast Fourier transform (FFT)
to compute some correlation function ($S_2, F_{ss}, F_{sv}, C_2$)
in all directions.
The FFT algorithm let us compute auto and cross-correlation fast,
especially on GPU.

@@

\section{The scan approach.}
\label{sec:scan}
In this approach, the input array is cut in parallel slices (``scanned'') along
each requested direction. Each slice is then analyzed independently as
one-dimensional system and the result is accumulated into \code{CorrelationData}
structure. In the following sections we provide the algorithms for computation
of the functions given in \cref{sec:math}.
